---
title: "Monte Carlo Simulations made easy and tidy with tidyMC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Monte Carlo Simulations made easy and tidy with tidyMC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# When to use tidyMC?

summarize typical tasks of MCs and how tidyMC helps to solve those: see [dpylr package ](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)


# Package overview

The `tidyMC` package aims to cover the whole workflow of running a Monte Carlo simulation in either an academic or professional setting. Thus, `tidyMC` aims to provide functions for the following steps:

* Running a Monte Carlo Simulation for a user defined function and given parameters using `future_mc()`
* Summarizing the results by (optionally) user defined summary functions using `summary.mc()`
* Creating plots of the Monte Carlo Simulation and its results, which can be modified by the user using `plot.mc()` and `plot.summary.mc()`
* Creating a LaTeX table summarizing the results of the Monte Carlo Simulation using `tidy_mc_latex()`

In the following subsections we will show you how you can implement those steps using the `tidyMC` package.


## Installing tidyMC

Until now, the `tidyMC` package is not on CRAN, thus you need to download the development version from [GitHub](https://github.com/stefanlinner/tidyMC) as follows:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("stefanlinner/tidyMC")
```

Afterwards you can load the package:

```{r}
library(tidyMC)
```


### Run your first Monte Carlo Simulation with `future_mc()`

`future_mc()` allows you to run a Monte Carlo Simulation for a user defined function and given parameters. The first argument of `future_mc()` is `fun` which has to be a function that handles the generation of data, the application of the method of interest and the evaluation of the result for a single repetition and parameter combination. `future_mc()` handles the generation of loops over the desired parameter grids and the repetition of the Monte Carlo experiment for each of the parameter constellations. Consider the following example for `fun` and note that it performs the required tasks of generating data, applying the method and evaluating the results: 

```{r}
# fun
ols_test <- 
  function(b0, b1, b2, n, sigma2, param_x1, param_x2, inc_x2){
    
    # generation of data
    x1 <- rnorm(n = n, mean = param_x1[1], sd = param_x1[2])
    x2 <- rnorm(n = n,  mean = param_x2[1], sd = param_x2[2])
    e <- rnorm(n, sd = sqrt(sigma2))
    y <- b0 + b1*x1 + b2*x2 + e
    
    if (inc_x2 == 0){
      x2 <- x2 * inc_x2
    }
    
    # application of method
    estim <- lm(y ~ x1 + x2)
    
    # evaluation of the result for a single repetition and parameter combination
    out <- list(b0 = estim$coefficients[1],
                b1 = estim$coefficients[2],
                b2 = estim$coefficients[3],
                sigma2 = var(estim$residuals))
    return(out)
  }
```

The second argument of `future_mc()` is `repetitions` which should be an integer specifying the number of Monte Carlo iterations. While the third argument `param_list` should be a list whose components are named after the parameters of `fun` which should vary for the different Monte Carlo Simulation and each component is a vector containing the desired grid values for the parameter. Consider the following example for `param_list` and note that its components are named accordingly to the parameters of `ols_test`: `n` and `inc_x2`, respectively:

```{r}
# param_list
param_list_ols <- 
  list(n = c(100, 200, 300), inc_x2 = c(0,1))
```

`future_mc()` takes care of creating all possible parameter combinations of `param_list` and runs the Monte Carlo Simulation for all possible combinations. The `...` argument can be used to specify further arguments of `fun` which are not contained in `param_list`. Those arguments will be held fixed for all parameter combinations. In our case those arguments are `b0`, `b1`, `b2`, `sigma2`, `param_x1`, and `param_x2`.

Moreover, there are four formal requirements that `fun` and thus `ols_test` have to fulfill:

* The arguments of `fun` which are present in `param_list` have to be scalar values. Note that the arguments of `ols_test` which are contained in `param_list_ols`: `n` and `inc_x2` are scalar values. The remaining arguments of `ols_test` are allowed to take non-scalar values.
* Every variable used inside `fun` has either to be defined inside `fun` or given as an argument through the `...` argument.
* The value returned by `fun` has to be a named list. In our example the names of the returned list are `b0`, `b1`, `b2`, and `sigma2`.
* The names of the returned values and those of the arguments contained in `param_list` need to be different. Moreover, they cannot be `params`, `repetitions` or `setup` as these names are already occupied. Note that `b0`, `b1`, `b2`, and `sigma2` are the names of the returned values as well as names of arguments of `ols_test`. However, none of those arguments is contained in `param_list_ols`. If we would add either of those variables to `param_list_ols` we would need to change the name of the returned value for the respective variable.

We recommend to even further restrict the return value of `fun` to be a named list of scalars. This allows you to use all comfort functions of the `tidyMC` package. As you can see, we did that for `ols_test`.

The argument `parallelisation_plan` allows the user to set a parallelisation plan. While the argument `parallelisation_options` allows the user to fine tune functions, such as `furrr::future_map()` by `furrr::furrr_options()`. Moreover, the user can also decide not to run the Monte Carlo in parallel by setting `parallel = FALSE`. As default `future_mc()` runs a quick check by running a single test-iteration for each parameter combination in order to check for possible errors in `fun`. If a error occurs the user not only receives the error message but also the parameter combinations for which the error occurred:

```{r, error=TRUE}
set.seed(101)
first_mc_ols <- future_mc(
  fun = ols_test, 
  repetitions = 1000, 
  param_list = param_list_ols, 
  b0 = 1, 
  b1 = 4, 
  b2 = 5, 
  sigma2 = -2,
  param_x1 = c(0,5),
  param_x2 = c(0,6),
  check = TRUE
)
```

The attentive reader might already have noticed that we specified `sigma2 = -2` which doesn't make sense, as the variance of the error term can not be negative. This results in a failed check for all parameter combinations, as this parameter is held fixed for any combination. Once we correct that mistake, we can run our first Monte Carlo Simulation:

```{r}
set.seed(101)
first_mc_ols <- future_mc(
  fun = ols_test, 
  repetitions = 1000, 
  param_list = param_list_ols, 
  b0 = 1, 
  b1 = 4, 
  b2 = 5, 
  sigma2 = 2, # correctly specify sigma2
  param_x1 = c(0,5),
  param_x2 = c(0,6),
  check = TRUE
)
```

`future_mc` returns a list of type `mc` and length `r length(first_mc_ols)` consisting of a tibble (`first_mc_ols$output`) containing the return value of `fun` for each iteration and parameter combination. In our case `first_mc_ols$output` contains a column for each output `b0`, `b1`, `b2`, and `sigma2`, as well as a column for each parameter in `param_list_ols` and a column containing the `nice_names` of the parameter combinations. Overall the `first_mc_ols$output` consists of `r nrow(first_mc_ols$output)` rows, i.e., for each parameter combination 1.000 rows: 

```{r}
first_mc_ols$output
```

If `ols_test` would not return a named list of scalars, but a named list of non-scalars, then `first_mc_ols$output` would not contain a column for each output, but a single column containing the named list of non-scalars for each iteration and parameter combination.

Moreover, `first_mc_ols` returns much other information about the Monte Carlo Simulation that can be printed in a dense representation:

```{r}
first_mc_ols
```


### Summarize your results with `summary.mc()`

If `fun` was further restricted to return a named list of scalars the function `summary.mc()` can be used to summarize the Monte Carlo results of an object of class `mc` that is returned by `future_mc()`. Using the argument `sum_funs` the user can define (different) functions which summarize the simulation results for each output (return values of `fun`) and each parameter combination. Thus, the functions inside `sum_funs` only take one argument, which is the output vector (with length `repetitions`) of one output of one specific parameter combination. The default summary functions are `mean()` for numeric outputs and `summary()` for outputs with non-numeric data types. As `ols_test` outputs are all numeric `mean()` will be used to summarize all outputs by default:

```{r}
# Default
summary_default <- summary(first_mc_ols)
summary_default
```

The user can define summary functions by supplying a named (nested) list to `sum_funs`. When the functions provided for each output return only one numeric value (like `mean()` but unlike `summary()`) the results are twofold: first, a single scalar result of the function evaluating the whole output vector. Second, a "path" with length `repetitions` of the step wise calculation of the function's result across the output vector. Thus, for `ols_test`

If the user wants to summarize the simulation results of a respective output in the same way for each parameter combination, a list whose components are named after the outputs (names of the returned list of `fun`) is supplied and each component is a function which only takes the vector of results of one output as the main argument.

If the user wants to summarize the simulation results of a respective output differently for different parameter combinations, a nested list has to be supplied. The components of the outter list must be equal in length and naming to the `nice_names` (see `first_mcs$nice_names`) and each component is another list (inner list). The components of the inner list are then defined the same way as above (components named after the outputs and each component is a function).

Here an example for the different specifications of `sum_funs`:

```{r}
# summarizing output the same way for each parameter combination with own functions
## summary functions only have one argument
quantile_sum <- function(x) quantile(x, probs = 0.75)
sum_funs1 <- list(b0 = quantile_sum, b1 = sd, b2 = mean, sigma2 = median)

summary_output_spec <- summary(first_mc_ols, sum_funs = sum_funs1)

# summarizing output differently for different parameter combinations
sum_funs2 <-
  list(
    list(b0 = min, b1 = min, b2 = min, sigma2 = mean),
    list(b0 = mean, b1 = mean, b2 = mean, sigma2 = mean),
    list(b0 = median, b1 = median, b2 = median, sigma2 = mean),
    list(b0 = max, b1 = max, b2 = max, sigma2 = mean),
    list(b0 = min, b1 = min, b2 = min, sigma2 = mean),
    list(b0 = mean, b1 = mean, b2 = mean, sigma2 = mean)
  )
names(sum_funs2) <- first_mc_ols$nice_names
summary_out_param_spec <- summary(first_mc_ols, sum_funs = sum_funs2)

```

The provided summary functions are not restricted regarding the complexity of their return value. However, the path of the summarized output over all simulation repetitions is only returned if the provided summary functions return a single numeric value. Thus, the following comfort functions `plot.summary.mc()` and `tidy_mc_latex()` will only work in this specific case.


### Plot your Monte Carlo Simulation with `plot.mc()` and `plot.summary.mc()`







### Create a LaTeX table of your results with `tidy_mc_latex()`
